|

### 2022-08-22 manuscript_BE/manuscript_BE.Rmd
Implementation of a probability for choosing the agent type is successfully implemented. However, the new model does not add a lot of explanatory power

```{=latex}
\keywords{learning, behavior, cleaners, bayesian statisitics, behavioral mechanisms}
```

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
($62cm \times 27cm \times 37 cm$) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to decide based solely on the association between 
the behaviour and the collor/pattern cue [@wismer_Cuebased_2019]. 
Triki et al. [-@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reached a learning criterion, 
that is, performing significantly above chance level in a binomial test 
($p-value 	\leq 0.05$, ),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped. The reversal phased 
stopped when the fish performed significatly above chance, or the fish
completed the 200 trial together with the initial;
see Triki et al.[-@triki_Biological_2019;@triki_Brain_2020]. 
Here, we are interested in explaining the total frequency of visitor choices
using the model, rather than just the achievement of the criterion. 
Total frequency of visitor choices naturally comes out of the model, 
and allows us to use all the variation among cleaners, instead of reducing 
that to a binomial variable.
Thus, we used instead a subset of these data to estimate the final 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two sessions (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trial-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish).

@@

source(here("analysisMCMC.R"))

ggarrange(allGammapost.plot,alletapost.plot,common.legend = TRUE)
# first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
#           gam.both.post.FAA,gam.gam.post.FAA,nrew.nrew.post.FAA,
#           labels=c('a','c','d'))
# second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
#           nrew.both.post.FAA,panel.rsqr.all,rel_heights = c(1,2),
#           labels = c('b','e'))
# plot_grid(ncol = 2,first.col,second.col)

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centered around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models' fit 
favors the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though the peak of the three
$pseudo-R^2$ distributions were not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of the black line in Fig. \ref{fig:post} e). This
shows that accounting for variation in the parameter estimates the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

@@

```{r pred, fig.cap='Observed and predicted probability of choosing a visitor. Left-hand side panel: colour contour shows the prediction from the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 reef sites, as well as the corresponding relative cleaner fish abundance (x axis) and frequency of visitors clients leaving the cleaning station (y axis). Right-hand side panels: Variation of the predicted probabilities  of choosing a visitor over a resident and their observed values for 12 locations. Circles show the mean prediction for each location from 100 samples taken from the posterior distribution. Thick and thin bars show the 66 and 95\\% credible interval, respectively, taken from those posterior samples. Squares show the predictions used for the panel on the right-hand side. Colour coding denotes different reef site/year of the data collection (see Empirical data section). The black line corresponds to a perfect match between observed and predicted probabilities. Upper panels (a and b) show predictions from a model including chaining and penalty; middle panels (c and d) from a  model with only chaining; lower panels (e and f) from a model with penalty only.',out.width='100%',fig.height=5,cache=FALSE}

 source(here("predictions.R"))
first_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          cont.obs.pred.both,cont.obs.pred.gam,cont.obs.pred.Nrew,
          labels=c('a','c','e'),rel_heights = c(1.4,1,1))
second_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          scatter.obs.pred.both,scatter.obs.pred.gam,scatter.obs.pred.Nrew,
          labels=c('b','d','f'))
plot_grid(first_col,second_col,rows = 1)
```

@@

sca.both.post.PAA<-ggplot(data=MCMCdata4,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.4$scaleConst))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,show.legend=FALSE) + 
  labs(title=labelsPlot.Scen[1],x="",y="")+
  theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))
sca.gam.post.PAA<-ggplot(data=MCMCdata5,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.5$scaleConst))),
               point_interval = mode_hdi, 
               .width = c(.66, .95),point_size=3,
               show.legend=FALSE)+theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[2],x="",y="")
sca.nRew.post.PAA<-ggplot(data=MCMCdata6,aes(x=scaleConst)) +
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.6$scaleConst))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("skyblue","gray85"))+
  labs(title=labelsPlot.Scen[3],x="",y="")+
  theme_classic()
ggarrange(sca.both.post.PAA,sca.gam.post.PAA,
          sca.nRew.post.PAA,
          labels=c('a','b','c'),common.legend = F)

@@

ggarrange(traces.neg.PAA,gelman.neg.PAA)
```

>>>>

<!-- ```{=latex} -->
<!-- \keywords{learning, behavior, cleaners, bayesian statisitics, behavioral mechanisms} -->
<!-- ``` -->

@@

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
($62cm \times 27cm \times 37 cm$) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to decide based solely on the association between 
the behaviour and the color/pattern cue [@wismer_Cuebased_2019]. 
Triki et al. [-@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reached a learning criterion, 
that is, performing significantly above chance level in a binomial test 
($p-value 	\leq 0.05$, ),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped. The reversal phased 
stopped when the fish performed significatly above chance, or the fish
completed the 200 trial together with the initial;
see Triki et al.[-@triki_Biological_2019;@triki_Brain_2020]. 
Here, we are interested in explaining the total frequency of visitor choices
using the model, rather than just the achievement of the criterion. 
Total frequency of visitor choices naturally comes out of the model, 
and allows us to use all the variation among cleaners, instead of reducing 
that to a binomial variable.
Thus, we used instead a subset of these data to estimate the final 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two sessions (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trial-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish).

@@

scen1<-"MCMCclean_gam_Nrew_sca_"
scen2<-"MCMCclean_gam_sca_"
scen3<-"MCMCclean_Nrew_sca_"
scen4<-"MCMCclean_PAA_gam_Nrew_sca_"
scen5<-"MCMCclean_PAA_gam_sca_"
scen6<-"MCMCclean_PAA_Nrew_sca_"

source(here("analysisMCMC.R"))
ggarrange(allGammapost.plot,alletapost.plot,common.legend = TRUE)
# first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
#           gam.both.post.FAA,gam.gam.post.FAA,nrew.nrew.post.FAA,
#           labels=c('a','c','d'))
# second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
#           nrew.both.post.FAA,panel.rsqr.all,rel_heights = c(1,2),
#           labels = c('b','e'))
# plot_grid(ncol = 2,first.col,second.col)

@@

```{r fitAllModels}
panel.rsqr.all
```

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centered around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models' fit 
favors the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though the peak of the three
$pseudo-R^2$ distributions were not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of the black line in Fig. \ref{fig:post} e). This
shows that accounting for variation in the parameter estimates the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConstFAA} and \ref{fig:scaConstPAA}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfullFAA},\ref{fig:diaggamFAA},\ref{fig:diagNegFAA},
\ref{fig:diagnosticsfullPAA},\ref{fig:diaggamPAA},\ref{fig:diagNegPAA}).

@@

```{r predFAA, fig.cap='Observed and predicted probability of choosing a visitor. Left-hand side panel: colour contour shows the prediction from the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 reef sites, as well as the corresponding relative cleaner fish abundance (x axis) and frequency of visitors clients leaving the cleaning station (y axis). Right-hand side panels: Variation of the predicted probabilities  of choosing a visitor over a resident and their observed values for 12 locations. Circles show the mean prediction for each location from 100 samples taken from the posterior distribution. Thick and thin bars show the 66 and 95\\% credible interval, respectively, taken from those posterior samples. Squares show the predictions used for the panel on the right-hand side. Colour coding denotes different reef site/year of the data collection (see Empirical data section). The black line corresponds to a perfect match between observed and predicted probabilities. Upper panels (a and b) show predictions from a model including chaining and penalty; middle panels (c and d) from a  model with only chaining; lower panels (e and f) from a model with penalty only.',out.width='100%',fig.height=5,cache=FALSE}

scen1<-"MCMCclean_gam_Nrew_sca_"
scen2<-"MCMCclean_gam_sca_"
scen3<-"MCMCclean_Nrew_sca_"
 source(here("predictions.R"))
first_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          cont.obs.pred.both,cont.obs.pred.gam,cont.obs.pred.Nrew,
          labels=c('a','c','e'),rel_heights = c(1.4,1,1))
second_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          scatter.obs.pred.both,scatter.obs.pred.gam,scatter.obs.pred.Nrew,
           rel_heights = c(1.4,1,1),
          labels=c('b','d','f'))
plot_grid(first_col,second_col,rows = 1)
```

```{r predPAA, fig.cap='Observed and predicted probability of choosing a visitor. Left-hand side panel: colour contour shows the prediction from the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 reef sites, as well as the corresponding relative cleaner fish abundance (x axis) and frequency of visitors clients leaving the cleaning station (y axis). Right-hand side panels: Variation of the predicted probabilities  of choosing a visitor over a resident and their observed values for 12 locations. Circles show the mean prediction for each location from 100 samples taken from the posterior distribution. Thick and thin bars show the 66 and 95\\% credible interval, respectively, taken from those posterior samples. Squares show the predictions used for the panel on the right-hand side. Colour coding denotes different reef site/year of the data collection (see Empirical data section). The black line corresponds to a perfect match between observed and predicted probabilities. Upper panels (a and b) show predictions from a model including chaining and penalty; middle panels (c and d) from a  model with only chaining; lower panels (e and f) from a model with penalty only.',out.width='100%',fig.height=5,cache=FALSE}

scen1<-"MCMCclean_PAA2_gam_Nrew_sca_"
scen2<-"MCMCclean_PAA2_gam_sca_"
scen3<-"MCMCclean_PAA2_Nrew_sca_"

source(here("predictions.R"))
first_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
          cont.obs.pred.both,cont.obs.pred.gam,cont.obs.pred.Nrew,
          labels=c('a','c','e'),rel_heights = c(1.4,1,1))
second_col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
                      rel_heights = c(1.4,1,1),
          scatter.obs.pred.both,scatter.obs.pred.gam,scatter.obs.pred.Nrew,
          labels=c('b','d','f'))
plot_grid(first_col,second_col,rows = 1)
```

@@

sca.both.post.PAA<-ggplot(data=MCMCdata4,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(level)),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,show.legend=FALSE) + 
  labs(title=labelsPlot.Scen[1],x="",y="")+
  theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))
sca.gam.post.PAA<-ggplot(data=MCMCdata5,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(level)),
               point_interval = mode_hdi, 
               .width = c(.66, .95),point_size=3,
               show.legend=FALSE)+theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[2],x="",y="")
sca.nRew.post.PAA<-ggplot(data=MCMCdata6,aes(x=scaleConst)) +
  stat_halfeye(aes(fill=stat(level)),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,
               show.legend=FALSE) +
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[3],x="",y="")+
  theme_classic()
ggarrange(sca.both.post.PAA,sca.gam.post.PAA,
          sca.nRew.post.PAA,
          labels=c('a','b','c'),common.legend = F)

@@

ggarrange(traces.neg.PAA,gelman.neg.PAA)
```
```{r diagbothGroup,fig.cap='MCMC convergence diagnostics for the penalty model. On the left trace-plots, on the right changes along the chain of the Gelman and Rubin shrink factor',out.width='100%',fig.height=5,cache=FALSE,message=FALSE,out.width='100%',fig.height=5,cache=FALSE,message=FALSE}
longMCMCdata7<-melt(MCMCdata7,id.vars = c("iteration","seed"),
                    variable.name = "parameter",
                    measure.vars = c("gamma","probFAA","probFAA.1","scaleConst"))

longMCMCdata7[,parameter:=factor(parameter,labels = c("gamma","p[FAA]","p[FAA_1]","'Scaling constant'"))]

traces.both.group<-ggplot(longMCMCdata7,aes(x=iteration,y=value,colour=as.factor(seed)))+
  geom_line()+
  scale_color_manual(values = multDiscrPallet)+
  labs(color="Chain")+
  facet_grid(rows=vars(parameter),scales = "free",switch = "y",
             labeller = label_parsed)+
  theme_classic()+
  theme(legend.position = 'top')

pdf(file = NULL)
gp.dat.both.group<-gelman.plot(mcmcList.7)
rubbish <- dev.off()

gp.dat.frame.both.group <- data.table(rbind(as.data.frame(gp.dat.both.group[["shrink"]][,,1]),
                          as.data.frame(gp.dat.both.group[["shrink"]][,,2])),
                q=rep(dimnames(gp.dat.both.group[["shrink"]])[[3]], each=nrow(gp.dat.neg[["shrink"]][,,1])),
                last.iter=rep(gp.dat.both.group[["last.iter"]], length(gp.dat.both.group)))

gelman.both.group<-ggplot(melt(gp.dat.frame.both.group, c("q","last.iter"), value.name="shrink_factor"),
       aes(last.iter, shrink_factor, colour=q, linetype=q)) +
  geom_hline(yintercept=1, colour="grey30", lwd=0.2) +
  geom_line() +
  facet_grid(rows = vars(variable)) +
  labs(x="Last Iteration in Chain", y="Shrink Factor",
       colour="Quantile", linetype="Quantile") +
  scale_linetype_manual(values=c(2,1))+
  theme_classic()+
  theme(legend.position = 'top',strip.text.y = element_blank())

ggarrange(traces.both.group,gelman.both.group)
```
